---
title: "WRAPPING UP"
subtitle: "and final projects"
author: "Austin Hart"
institute: "American University"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, rladies, rladies-fonts]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  message=FALSE, warning=FALSE, eval = TRUE, echo = FALSE, 
  fig.align = 'center', dev='svg'
)
```


```{r results='hide'}
library(tidyverse)
library(magrittr)
library(kableExtra)

```

# Plan for today

- Overview of course content

- Final project prep
  - Setting your schedule
  - Building from presentation

- Open for questions


---
class: middle, right, inverse

# COURSE IN REVIEW
### what did we even do?


---
# The course in brief

.left-column[

### our goal


]

.right-column[

- Ask and answer empirical questions.

- Assess and critique causal claims.

- Participate in and contribute to the advancement of human knowledge.

]

---
# The course in brief

.left-column[

### our goal

### framework

]

.right-column[

- Modern scientific approach
  - ask questions
  - formalize hypotheses
  - collect data
  - reject the implausible

- Quantitative statistical analysis
  - Large-N approach
  - Descriptive analysis
  - Null hypotheses testing
  - Regression/Multiple regression

]


---
# The course in brief

.left-column[

### our goal

### framework

### why bother?

]

.right-column[

> THE METHOD MATTERS

- Fundamental Problem of Causal Inference
  - Effect of T (vs C) on Y $= Y_{i,T} - Y_{i,C}$
  - But can only see $i$ exposed to $T$ or $C$.
  - We can estimate but **never observe** causal effect
  
- Hurdles to valid causal inference 
  - Evidence of empirical association
  - Causal mechanism
  - Eliminate bias from confounding
  - Eliminate bias from reverse causality

]



---
# The course in brief

.left-column[

### our goal

### framework

### why bother?

### role of stats

]

.right-column[

Where do "the stats" fit in?  

- Describe the world
  - Summary stats and visualization
  - Frequency/counts

- Establish empirical association
  - Difference of means
  - Regression analysis
  
- Eliminate confounding bias
  - Multiple regression
  - Experimental design

]


---
class: inverse, middle, right

# CORE CONCEPTS



---
# Big ideas to know

- Causal inference and FPCI

- Experimental vs observational designs

- Hurdles to valid causal inference

- Hypothesis testing and p-values

- Multiple regression and confounding bias


---
class: inverse, right, middle

# CORE SKILLS
### and when to use them


---
# Core skills

.left-column[

### Summary stats

]


.right-column[

To summarize data on a **numeric** variable:  

- Central tendency
  - Mean: first choice; biased in skewed dist
  - Median: safe option in presence of skew
  
- Dispersion
  - Standard deviation: pair with mean
  - Interquartile range (IQR): pair with median
  
- Visualization
  - Histogram
  - Boxplot

]



---
# Core skills

.left-column[

### Summary stats

### Freqs/counts

]


.right-column[

To summarize data on a **categorical** variable:  

- Presentation
  - Relative frequency table
  - Bar chart
  
- Critical values
  - Modal category
  - Modal frequency

]




---
# Core skills

.left-column[

### Summary stats

### Freqs/counts

### Diff means

]


.right-column[

Difference of means   

- Evaluate association between
  - **numeric** outcome variable
  - **binary** exposure variable
  
- Presentation
  - Box plot or table of group means
  - Write out and compare group means
  
- Hypothesis testing, $t-$test
  - Null: no mean difference $\mu_T - \mu_C = 0$
  - p-value: $Pr[(Y_T - Y_C) ~ | ~ (\mu_T - \mu_C = 0)]$

]


---
# Core skills

.left-column[

### Summary stats

### Freqs/counts

### Diff means

### Regression

]


.right-column[

Regression analysis  

- Evaluate association between
  - **numeric** outcome variable
  - **numeric** exposure variable
  
- Presentation
  - Scatter plot with linear fit
  - Regression table
  - Interpret coefficient estimates
  
- Hypothesis testing
  - Null: $\beta = 0$
  - p-value $= Pr(b_X ~ | ~ \beta_X=0)$

]


---
class: inverse, middle, right

# FINAL PROJECT

### beyond the presentation


---
# Final project

.left-column[

### What and when?

]


.right-column[

- Deliverables
  - Written analysis memo (1000-1500 words)
  - Replication dataset ('LastName.Rdata')
  - Replication script ('LastName.R')

- DUE TUESDAY, DECEMBER 13

]


---
# Final project

.left-column[

### What and when?

### The memo

]


.right-column[

- Executive summary
- Background and expectation
- Data
  - Source, coverage, measurement
  - Graph and description of outcome var
- Analysis
  - Graph of bivariate relationship
  - Table of bivariate and multiple regression
  - Clear interpretation
- Conclusion

]


---
# Final project

.left-column[

### What and when?

### The memo

### Prep

]


.right-column[

- Build from feedback
- Follow the guidelines
- Care in format on tables and graphs
- Clear question
- Clear interpretation of tests and results

]




---
# The final exam

.left-column[

### What and when?

### The memo

### Prep

### Beware

]


.right-column[

- Give yourself TIME!
  - **Don't have data?** Red alert!
  - **Fix/improve analysis?** Do this first.
  - **Writing and interpretation** Think 2 or 3 drafts

- Work must be your own.
  - Your code.
  - Your visuals.
  - Your analysis.

]



---
class: inverse

# Wrapping up

### Importance of a scientific approach

- Asking questions

- Challenging convention

- Building knowledge

